# Image Classifier Project
# CIS-544-11
# Cat and Dog dataset

# To load the required libraries
```{r}
require(h2o)
require(tidyverse)
require(h2o)
require(caTools)
require(naivebayes)
require(randomForest)
require(rvest)
require(imager)
```

# To demonstrate imager package on one of the files

```{r}
sample_image <- load.image("C:/Users/s0853301/Desktop/fp/cat/cat.1.jpg")
```

```{r}
# Edge detector for x axis
sample.image.xedges <- deriche(sample_image,2,order=2,axis="x")
plot(sample.image.xedges)
```


```{r}
# Edge detector for y axis
sample.image.yedges <- deriche(sample_image,2,order=2,axis="y")
plot(sample.image.yedges)
```

```{r}
#Chain operations using the pipe operator (from magrittr)
deriche(sample_image,2,order=2,axis="x") %>% deriche(2,order=2,axis="y") %>% plot
```


```{r}
#Another example of chaining: image gradient along x and y axes
layout(matrix(1:2,1,2));
grayscale(sample_image) %>% imgradient("xy") %>% plot(layout="row")
```


# To define a function which generates features for a given image and then loop through the entire list of images 

```{r}
convertonumbers <- function(image,label){
# Load image 
im <- load.image(image)
# Detect edges and set standard size
im <- deriche(im,2,order=2,axis="x") %>% deriche(2,order=2,axis="y") %>% resize(50,50)
# Convert to Data Frame
df <- as.data.frame(im)
# Add Label
row <- data.frame(Label = label)
# Separate Colors
R <- subset(df,df$cc == 1)
G <- subset(df,df$cc == 2)
B <- subset(df,df$cc == 3)
# Create Grouping Vectors
c10 <- c(1:10)
c20 <- c(11:20)
c30 <- c(21:30)
c40 <- c(31:40)
c50 <- c(41:50)

# 3 Loops, one for each Color  of RGB

# Loop through X
for(i in 1:5) {
  x <- i
  # Change the Grouping Vector depending on x itteration
  onX <- case_when(x == 1 ~ c10,
                   x == 2 ~ c20,
                   x == 3 ~ c30,
                   x == 4 ~ c40,
                   x == 5 ~ c50)
  # Loop through y
  for (j in 1:5) {
    y <- j
    # Change the Grouping Vector depending on y itteration
    onY <- case_when(y == 1 ~ c10,
                     y == 2 ~ c20,
                     y == 3 ~ c30,
                     y == 4 ~ c40,
                     y == 5 ~ c50)
    # Change name of each column 
    valName <- paste0("x", i, "y", j, ".R")
    # Group based on the coordiate's groups
    temp <-
      R %>% filter(x %in% onX &
                     y %in% onY) %>% summarize(val = mean(value))
    colnames(temp) <- valName
    # Add to the row data frame
    row <- cbind(row, temp)
  }
  
}
# Repeat proces for G
for(i in 1:5) {
  x <- i
  onX <- case_when(x == 1 ~ c10,
                   x == 2 ~ c20,
                   x == 3 ~ c30,
                   x == 4 ~ c40,
                   x == 5 ~ c50)
  for (j in 1:5) {
    y <- j
    onY <- case_when(y == 1 ~ c10,
                     y == 2 ~ c20,
                     y == 3 ~ c30,
                     y == 4 ~ c40,
                     y == 5 ~ c50)
    valName <- paste0("x", i, "y", j, ".G")
    temp <-
      G %>% filter(x %in% onX &
                     y %in% onY) %>% summarize(val = mean(value))
    colnames(temp) <- valName
    row <- cbind(row, temp)
  }
  
}
# Repeat Process for B
for(i in 1:5) {
  x <- i
  onX <- case_when(x == 1 ~ c10,
                   x == 2 ~ c20,
                   x == 3 ~ c30,
                   x == 4 ~ c40,
                   x == 5 ~ c50)
  for (j in 1:5) {
    y <- j
    onY <- case_when(y == 1 ~ c10,
                     y == 2 ~ c20,
                     y == 3 ~ c30,
                     y == 4 ~ c40,
                     y == 5 ~ c50)
    valName <- paste0("x", i, "y", j, ".B")
    temp <-
      B %>% filter(x %in% onX &
                     y %in% onY) %>% summarize(val = mean(value))
    colnames(temp) <- valName
    row <- cbind(row, temp)
  }
  
}
# Return Final Data Set
return(row)
}

```


# Calling the above defined function

```{r}


# List the files of the Cars Folder
Cat_files <-
  list.files(path = "C:/Users/s0853301/Desktop/fp/Cat")
# Create Empty Data Frame
Cats <- data.frame()

setwd("C:/Users/s0853301/Desktop/fp")

# Loop each file
for (i in 1:length(Cat_files)) {
  # Get the path for each photo
  photo <- paste0("Cat/", Cat_files[i])
  # Use formula to generate the numbers and assign the label "Cat"
  data <- convertonumbers(photo, "Cat")
  # Bind Each Result together
  Cats <- rbind(Cats, data)
}


# Repeat process for the images that are not Cars
Dog_files <-
  list.files(path = "C:/Users/s0853301/Desktop/fp/Dog")
Dogs <- data.frame()

setwd("C:/Users/s0853301/Desktop/fp")

for (i in 1:length(Dog_files)) {
  photo <- paste0("Dog/", Dog_files[i])
  data <- convertonumbers(photo, "Dog")
  Dogs <- rbind(Dogs, data)
}
# Bind Cars and Not Cars data frames.
Data <- rbind(Cats, Dogs)
# Change the labels to factors
Data$Label <- as.factor(Data$Label)
```


# Setup the h2o environemtn and convert dataframe to an h2o object
```{r}
localh2o <- h2o.init()
SPLIT_RATIO = 0.75
h2o.cat.dog  <- as.h2o(Data)
set.seed(12345)
```

# Split the data
```{r}
h2oSplit <- h2o.splitFrame(data = h2o.cat.dog, ratios = SPLIT_RATIO)
train.h2o <- h2oSplit[[1]]
test.h2o <- h2oSplit[[2]]
```

# To build 4 separate classifiers 

# To build a NaiveBayes model 

```{r}
#NaiveBayes Model
nbModel <- h2o.naiveBayes(y = "Label",
                          training_frame = train.h2o,
                          validation_frame = test.h2o)

nbPredict <- h2o.predict(nbModel, test.h2o)

# Confusion Matrix for Naive Bayes

table.nb <- table(as.vector(nbPredict$predict), as.vector(test.h2o$Label)) 


nbresult <- data.frame(Actual = as.vector(test.h2o$Label), Prediction = as.vector(nbPredict$predict))

# TP,TN, FP, FN
TPnb <- subset(nbresult, Actual == "Cat" & Prediction == "Cat" )
TNnb <- subset(nbresult, Actual == "Dog" & Prediction == "Dog")
FPnb <- subset(nbresult, Actual == "Dog" & Prediction == "Cat")
FNnb <- subset(nbresult, Actual == "Cat" & Prediction == "Dog")

# Accuracy
accuracynb <- (nrow(TPnb)) / (nrow(TPnb) + nrow(FPnb))
cat("Naive Bayes Accuracy:" , accuracynb*100,"%", "\n")

# Precision

precision.nb <- (nrow(TPnb))/ (nrow(test.h2o))
cat("Naive Bayes Precision:" , precision.nb*100,"%", "\n")


# Recall
recallnb <- (nrow(TPnb)) / (nrow(TPnb) + nrow(FNnb))
cat("Naive Bayes Recall:" , recallnb*100,"%", "\n")

# Performance
nbPerformance <- h2o.performance(nbModel, test.h2o)

# Auc
h2onbAUC <- h2o.auc(nbPerformance)
cat("Naive Bayes AuC:", h2onbAUC*100,"%", "\n")
```

# To build a Random Forest model

```{r}
# RandomForest
rfModel <- h2o.randomForest(y = "Label",
                            training_frame = train.h2o,
                            validation_frame = test.h2o)
                            

rfPredict <- h2o.predict(rfModel, test.h2o, type = "class")

rfresult <- data.frame(Actual = as.vector(test.h2o$Label), Prediction = as.vector(rfPredict$predict))

# TP,TN, FP, FN
TPrf <- subset(rfresult, Actual == "Cat" & Prediction == "Cat" )
TNrf <- subset(rfresult, Actual == "Dog" & Prediction == "Dog")
FPrf <- subset(rfresult, Actual == "Dog" & Prediction == "Cat")
FNrf <- subset(rfresult, Actual == "Cat" & Prediction == "Dog")

# Accuracy
accuracyrf <- (nrow(TPrf)) / (nrow(TPrf) + nrow(FPrf))
cat("Random Forest Accuracy:" , accuracyrf*100,"%", "\n")

# Precision
precisionrf <- (nrow(TPrf)) / (nrow(test.h2o))
cat("Random Forest Precision:" , precisionrf*100,"%", "\n")

# Recall
recallrf <- (nrow(TPrf)) / (nrow(TPrf) + nrow(FNrf))
cat("Random Forest Recall:" , recallrf*100,"%", "\n")

# Performance
rfPerformance <- h2o.performance(rfModel, test.h2o)

# Auc
h2orfAUC <- h2o.auc(rfPerformance)
cat("Random Forest AuC:", h2orfAUC*100,"%", "\n")
```

# To build a Deep Learning model

```{r}
# DEEP LEARNING
h2oDL <- h2o.deeplearning(
  x = colnames(train.h2o),
  y = c("Label"),
  training_frame = train.h2o,
  hidden = c(30, 20, 10),
  epochs = 100,
  seed = 12345,
  nfolds = 3
)


DLpred <- h2o.predict(h2oDL, test.h2o)
DLresult <- data.frame(Actual = as.vector(test.h2o$Label), Prediction = as.vector(DLpred$predict))

# TP,TN, FP, FN
TPDL <- subset(DLresult, Actual == "Cat" & Prediction == "Cat" )
TNDL <- subset(DLresult, Actual == "Dog" & Prediction == "Dog")
FPDL <- subset(DLresult, Actual == "Dog" & Prediction == "Cat")
FNDL <- subset(DLresult, Actual == "Cat" & Prediction == "Dog")

# Accuracy
accuracyDL <- (nrow(TPDL)) / (nrow(TPDL) + nrow(FPDL))
cat("Deep Learning Accuracy:" , accuracyDL*100,"%", "\n")

# Precision
precisionDL <- (nrow(TPDL)) / (nrow(test.h2o))
cat("Deep Learning Precision:" , precisionDL*100,"%", "\n")

# Recall
recallDL <- (nrow(TPDL)) / (nrow(TPDL) + nrow(FNDL))
cat("Deep Learning Recall:" , recallDL*100,"%", "\n")

# Performance
DLPerformance <- h2o.performance(h2oDL, test.h2o)

# Auc
h2oDLAUC <- h2o.auc(DLPerformance)
cat("Deep Learning AuC:", h2oDLAUC*100,"%", "\n")
```

# To build a SVM model

# Conventional Split
```{r}
set.seed(12345)
sample <- sample.split(Data$Label, SplitRatio = 0.75)
train <- subset(Data, sample == T)
test <- subset(Data, sample == F)
```


```{r}
# Support Vector Machine
modelsvm <- ksvm(Label ~ ., data = train, kernel = "rbfdot")

predsvm <- predict(modelsvm,test)
svmresult <- data.frame(Actual = test$Label, Prediction = predsvm)

# TP,TN, FP, FN
TPsvm <- subset(svmresult, Actual == "Cat" & Prediction == "Cat" )
TNsvm <- subset(svmresult, Actual == "Dog" & Prediction == "Dog")
FPsvm <- subset(svmresult, Actual == "Dog" & Prediction == "Cat")
FNsvm <- subset(svmresult, Actual == "Cat" & Prediction == "Dog")

# Accuracy
accuracysvm <- (nrow(TPsvm)) / (nrow(TPsvm) + nrow(FPsvm))
cat("Gaussian RBF Kernel Accuracy:" , accuracysvm*100,"%", "\n")

# Precision
precisionsvm <- (nrow(TPsvm)) / (nrow(test.h2o))
cat("Gaussian RBF Kernel Precision:" , precisionsvm*100,"%", "\n")

# Recall
recallsvm <- (nrow(TPsvm)) / (nrow(TPsvm) + nrow(FNsvm))
cat("Gaussian RBF Kernel Recall:" , recallsvm*100,"%", "\n")


```
